\documentclass[a4paper, twoside]{report}

\input{packages.tex}

\title{A Deeper Dive into Relational Algebra by Way of Adjunctions}
\author{Matteo Bongiovanni}
\date{January 2023}
\supervisor{Dr.\ Nicolas Wu}
\secondmarker{Dr.\ Steffen van Bakel}
\documenttype{MEng Individual Project}
\university{Imperial College London}
\department{Department of Computing}
\degree{Joint Mathematics and Computing}

\begin{document}
\begin{comment}
General Information
The project report (occasionally referred to as a thesis) is an extremely
important aspect of the project. It serves to show what you have achieved and
should demonstrate that:

You understand the wider context of computing by relating your choice of project,
and the approach you take, to existing products or research.
You can apply the theoretical and practical techniques taught in the course to the
problem you are addressing and that you understand their relevance to the wider
world of computing.
You are capable of objectively criticising your own work and making constructive
suggestions for improvements or further work based on your experiences so far.
As a computing professional, you operate ethically and can explain your thinking
and working processes clearly and concisely to third parties who may not be
experts in the field in which you are working.
Most of the project assessors will not have followed the project throughout and
will only have a short time to listen to a presentation or see a demonstration.
For this reason they will rely heavily on the report to judge the project

Many students underestimate the importance of the report and make the mistake of
thinking that top marks can be achieved simply for producing a good product. This
is fundamentally not the case and many projects have been graded well below their
potential because of an indifferent or poor write-up. In order to get the balance
right you should consider that the aim of the project is to produce a good report
and that software, hardware, theory etc. that you developed during the project are
merely a means to this end. Don't make the mistake of leaving the write-up to the
last minute. Ideally you should produce the bulk of the report as you go along and
use the last week or two to bring it together into a coherent document.

It is helpful to get feedback from your supervisor about your project report, but
supervisors cannot be expected to look at documents written at the last minute or
at more than one chapter at a time. Allow plenty of time for this

Project Report Length
The nominal maximum report length is 60 pages, plus appendices. Shorter reports
may be appropriate for many projects and longer reports can be submitted with
prior approval (see below).

The page count includes:

title page
abstract
dedications
table of contents
The page count does not include:

bibliography
appendices (but the report should make sense without the appendices)
When writing up your work be reminded of the fact that a 60-page report isn't
necessarily better than a 30-page report, in the same way that a 10,000 line
implementation isn't necessarily better 5,000 line one. Conciseness, clarity and
elegance are invaluable qualities in report writing, just as they are in
programming. Indeed, you will lose credit if it is evident to the assessors that
the report has been padded with superfluous content in order to make it appear
more substantial than the work justifies. The assessors greatly value quality over
quantity.

It is important to understand that the report is assessed on the quality of the
technical writing. Here this means your ability to document your ideas, design,
implementation, evaluation, conclusions etc. succinctly, using appropriate
scientific language, as might a researcher when writing up their findings in an
technical magazine, academic journal or conference proceedings.

If you have good reason to use more than 60 pages you should consult your
supervisor and second marker and obtain their approval in advance - they will be
required to read and assess the whole report, whereas other assessment team
members may only scan key sections.

Project Layout
The physical layout and formatting of the report is also important. A tidy, well
laid out and consistently formatted document makes for easier reading and is
suggestive of a careful and professional attitude towards its preparation. Both
LaTeX and MS Word will allow you to produce a cleanly formatted document. Many
supervisors will advise you to use LaTeX as this solves most of the formatting
problems for you. If your report is to contain a substantial number of
mathematical formulae you are strongly advised to use LaTeX.

Project Report Structure
The exact structure of your project (the chapter titles etc) is up to you. Discuss
with your supervisor what structure works best for your project

A project report will usually contain a number of chapters (Chapter 1, Chapter 2
etc). Each chapter contains Sections (2.1, 2.2 etc) and may also contain
subsections (2.1.1, 2.1.2) etc. Try to avoid too many levels of subheading - three
is usually sufficient.

Projects will usually contain all of the following elements, although these may
not exactly map to chapters.
\end{comment}

% Front matter
\include{title/title}
\include{title/abstract}
\begin{comment}
Table of contents:
This should list the main chapters, sections and subsections of your report.
Choose self-explanatory chapter and section titles and use double spacing for
clarity. If possible you should include page numbers indicating where each chapter
section begins. 
\end{comment}
\tableofcontents

% Main matter 
\include{introduction/final}
\include{background/final}
\include{project/project}
\include{evaluation/evaluation}
\include{ethics/final}
\include{conclusion/conclusion}

% Back matter
\bibliographystyle{unsrtnat}
% DoC uses the Vancouver Referencing Format.
\bibliography{bibs/combined}

\end{document}
