\section{Benchmarking databases}
Databases are pervasive to modern society and thus standards have arisen over
the last few decades to ensure that customers are able to pick their preferred
DBMS vendor.

At the top level, database benchmarks are classified into three categories: industry-standard, vendor and
customer-application \cite{PractitionersIntroduction}. 
These classifications are usually motivated by intention of the benchmark
instead of structure of the database management system; there is no shortage of
papers emphasising the importance of domain-specific benchmarks for applications
\cite{PractitionersIntroduction, BenchmarkHandbook} and depending on the
risk/performance tolerance of the application it may be necessary to consider
results from all three categories.

\paragraph{Vendor benchmark} A vendor database benchmark is used by
the database vendor during the production of the database management system.
This benchmark usually serves multiple purposes not just limited to the design
of the system. Of course, it is often used to highlight any performance
bottlenecks driving the design of the system internally but it usually doubles
up and acts as the knowledge base for the marketing of the system. Vendor
benchmarks are usually characterised by a more comprehensive suite of
tests to generate the insights capable of driving the direction of the
product \cite{PractitionersIntroduction}.

\paragraph{Industry-standard benchmark} An industry-standard benchmark is a set
benchmarking suite designed independently to any vendor or solution. It is
designed to allow for a fair comparison between different vendors and has been
shown to increase competition between vendors \cite{Wisconsin2}. As much of
this review will show, many industry-standard benchmarks have been developed
to give results for a wide range of applications of databases for much more
relevant and specialised results; and, although many benchmarks have similar
metrics \cite{SetQueryBenchmark, DebitCredit}, just selecting what metrics to
show consumers is not a trivial task \cite{DebitCredit}. When Gray's paper
\cite{BenchmarkHandbook} was written, it was noted that these benchmarks were
becoming so popular that vendors were also beginning to report their results
with their marketing.

\paragraph{Consumer-application benchmark} This type of benchmark refers to any
benchmarking that a customer would run, typically to choose between different
vendors for their application. This kind of testing can be critical for a
performance-sensitive application \cite{PractitionersIntroduction} and is often
done to test the performance of the database under a specific installation
profile \cite{DoingYourOwnBenchmark} or loads. This is a very specialised
requirements based benchmarking.

The review now turns to different domain-specific benchmarking of databases in
order to comment on and analyse any common structures found when designing
databases. This will help inform major design decisions of this project when
deciding what aspects to include in the benchmarking of the alternative join
syntax.

Despite the innumerable use cases for databases in the modern world, many
applications require and prioritise similar values when choosing a solution.
Furthermore, as most industrial-standard database benchmarks have been designed in
order to allow comparison for specific domain requirements, you will commonly
find database benchmarks that are designed with one of two different
applications in mind: \emph{transaction processing} or \emph{decision support}
\cite{PractitionersIntroduction}; it is worth mentioning that other application
types exist, however, such as \emph{document search} and \emph{direct marketing}
\cite{SetQueryBenchmark}. The application types of transaction processing and
decision support are so pervasive that they are commonly also used to partition
other application types, for instance OLTP (Online transaction processing) and
OLAP (Online analytical processing) are specialised for online processing but
share many comparative similarities as transaction processing and decision
support respectively \cite{OLTP-Oracle}.

This review will provide a brief overview of both
the main database application types, then focus more strongly on decision
support benchmarks as ad-hoc query processing is more heavily tested in these
benchmarks and thus more relevant to the project.

\subsection{Transaction processing}
Transaction processing is characterised by a large number of
update-intensive database services with a particular emphasis on integrity of
its requests \cite{PractitionersIntroduction}.
It is clear that this type of environment
demands an emphasis on throughput and integrity. For instance, a bank would
likely have an online transaction processing system in place and the
ability to deal with a large number of transactions in a short period of
time while maintaining the accuracy of the information is most likely paramount.
There are many well known benchmarks to test these types of systems,
including \emph{DebitCredit}, \emph{TPC-C} and \emph{TPC-E} \cite{DebitCredit,
TPC-OLTP}.
These benchmarks usually measure transactions per second or price per
transactions per second. Efforts have been made to try and describe what is
meant by ``price'' and usually it refers to the five-year capital cost (e.g.
cost of hardware, software and system maintenance over a five year period)
\cite{BenchmarkHandbook, DebitCredit}. The advantage of giving a metric
dependent on price is that price is likely to be a limiting of important factor
to most (if not all) organisations when shopping for database solutions and so
it helps encompass all information about the system into a single comparable
number \cite{SetQueryBenchmark}. It is worth noting that many benchmarks which
include this metric also give more information for the user \cite{TPC-H} or
allow the user to weight the comparative importance of features in this metric
\cite{SetQueryBenchmark} but it is expected that much emphasis will come down to
this defining number \cite{SetQueryBenchmark}.

We note that the above definition of a transaction processing system not only
highlights its intense performance requirements but also notes the importance of
integrity. The bank example highlights intuitively what we mean by integrity, a
shop may not want the transaction to go through zero times if it is reported
that it has, in fact, been processed, just as I would not want the transaction
to go through twice. So what exactly is a transaction? And how can we specify
its integrity in a rigorous way?

A transaction simply is an input to the database that must be considered as one
unit of work \cite{ComputerScienceDictionary} and completed independently to any
concurrent actions occurring \cite{ACID}. The integrity of this behaviour can be described 
by a set of properties typically remembered as the acronym \emph{ACID} and
explained below \cite{ACID, PractitionersIntroduction}.

\paragraph{Atomicity} An ``all or nothing'' property of a transaction. All
individual operations must take place or any partially-completed sequence of
operations must have no lasting effects on the data. The user should be informed
of which outcome has occurred.

\paragraph{Consistency} Given the database was in a consistent/valid state
before the transaction processed, it must remain in a consistent/valid state
after any changes are committed; that is to say the changes cause not integrity
constraints to be violated \cite{IntroToDatabaseSystems}. 
Some texts favour the term ``correctness'' \cite{IntroToDatabaseSystems} but this
discussion is outside of the scope of the project. 
% correctness vs consistency: correctness may be preferred as it is stronger, 
% and says that the database only contains true statements about the world

\paragraph{Isolation} Information related to the execution of one transaction
should be hidden from other transactions running concurrently. This is
equivalent to being able to find an order such that the results of the execution
of all transactions would be the same if they were executed sequentially in that
order.

\paragraph{Durability} Once a transaction has been successfully completed, its
effects must persist and future malfunctions.

\subsection{Decision support}
Decision support is characterised by the need to execute complicated queries on
a database with fewer changes \cite{IntroToDatabaseSystems,
PractitionersIntroduction}. The business value from this sort of application are
vast, especially in automatic report generation to produce insights and direct
company strategy \cite{SetQueryBenchmark, IntroToDatabaseSystems}.
% Mention Wisconsin benchmark, single-user decision support and set query
% benchmark as examples of benchmarks

\subsection{Database benchmarking case studies}
% Explain why selected these benchmarks

% Wisconsin
% Set query
% DebitCredit

\subsection{Best practices benchmarking}
