\section{Evolution of database representation}\label{sec:background:dbrep}
This section introduces the mathematical framework presented in \relalg{} for
defining the relational model. It is useful to help understand the structure in
\fref{chap:database}.

\subsection{Bags}
\paragraph{Characteristics of a database} We expect our database approximation to not be ordered and admit multiplicities and a finite bag of values is one of the simplest constructions that does so. Like a finite set, a bag contains a collection of unordered values. However, unlike a set, bags can contain duplicate elements.  This multiplicity is key for processing non-idempotent aggregations. For instance, if summing up the ages of a database of people, without admitting multiplicity we would only sum each unique age once.

\subparagraph{Generalisation} Furthermore, going forward we generalise to bags
of any types instead of the classical ``bags of records''. This also allows us
to deal with intermediate tables that contain non-record values which, again,
may be useful for describing intermediate states of aggergations or projections.

In \fref{tab:BagRelAlgOps} we summarise the implementation of relational algebra operators with bags
as their bulk type \cite{RelationalAlgebraByWayOfAdjunctions}.
\begin{table}[h]
    \centering
    \begin{tabular}{r|l}
        table of $V$ values & \bag{V} \\
        empty table & \emptybag \\
        singleton table & \singletonbag \\
        union of tables & $\bagunion{}{}$ \\
        Cartesian product of tables & $\times$ \\
        neutral element & $\lbag () \rbag$ \\
        projection $\projsymb{f}$ & $\bag{f}$ \\
        selection $\selectsymb{p}$ & $filter\ p$ \\
        aggregation in monoid $\monoid{M}$ & $reduce\ \monoid{M}$\\
    \end{tabular}
    \caption{Relational algebra operators implemented for bags}
    \label{tab:BagRelAlgOps}
\end{table}

\subsection{Indexed tables}
We want to move towards an indexed representation of our table in order to
equijoin by indexing. In this section we introduce the mathematical concepts required to define such an implementation.
\theoremstyle{definition}\newtheorem*{psetdef}{Pointed set}
\theoremstyle{definition}\newtheorem*{ppfuncdef}{Point-preserving function}
\theoremstyle{definition}\newtheorem*{mapdef}{Map}
\theoremstyle{definition}\newtheorem*{finitemapdef}{Finite map}
\theoremstyle{definition}\newtheorem*{indexedtabledef}{Indexed Table}
\begin{psetdef}\label{def:pset}
  A pointed set $\pset{A}{a}$ is a set $A$ with a distinguished element $a \in A$.
\end{psetdef}
We commonly refer to the distinguished element of a set $A$ as $null_A$ or, when not ambiguous to do so, $null$.
\begin{ppfuncdef}\label{def:ppfunc}
  Given two pointed sets $\pset{A}{null_A}$ and $\pset{B}{null_B}$, a total function $f: A \rightarrow B$ is point-preserving if $f(null_A) = null_B$.
\end{ppfuncdef}

We now have the mathematical tools required to define a map. In its finite form a map is widely known in computer science by many other names such as a dictionary, association lists or key-value maps.

Let $\keyset$ be a set and $\valset$ a pointed set. To those already familiar with maps, it may help to think of $\keyset$ as keys and $\valset$ as values.
\begin{mapdef}
  A map of type $\map{\keyset}{\valset}$ is a total function from K to V.
\end{mapdef}
\begin{finitemapdef}
  A finite map of type \finitemap{\keyset}{\valset} is a map where only a finite number of keys are mapped to $null_\valset$ (where $null_\valset$ is the distinguished element of \valset). 
\end{finitemapdef}
The advantage of using a finite map in a database is to allow aggregation.
Furthermore, we note the following
isomorphisms~\cite{RelationalAlgebraByWayOfAdjunctions} where $1$ is the unit
type and pointed.

\begin{equation*}
\begin{split}
    & empty: 1 \rightarrow \finitemap{\keyset}{1} \\
    & empty = \lambda\:k \rightarrow () \\
    \\
    & merge: \finitemap{\keyset}{\valset_1} \times \finitemap{\keyset}{\valset_2}
             \rightarrow \finitemap{\keyset}{\valset_1 \times \valset_2} \\
    & merge (s, t) = \lambda\:k \rightarrow (s k, t k) \\
\end{split}
\end{equation*}

The functions above tell us some extremely important information on creating
empty maps and calculating their unions. As you can see $empty$ returns a
function that maps any key to the neutral element $()$. This is to be expected
as there are no values in an empty map. More interestingly, we see the merge
of two maps as a function that returns a pair of key lookups in each table.

We now have the correct machinery to define an indexed table.

\begin{indexedtabledef}
    An indexed table \indexedTable{\keyset}{\valset} is simply \finitemap{K}{\bag{V}}.
\end{indexedtabledef}

Set with the definitions above, \fref{tab:indexedTableRelAlgOps} can be referred
to for a summary of relational algebra operations implemented for a bulk type of
indexed tables~\cite{RelationalAlgebraByWayOfAdjunctions}.

\begin{table}[h]
    \centering
    \begin{tabular}{r|l}
        \keyset{}-indexed table of \valset{} values & \indexedTable{\keyset}{\valset} \\
        empty table & $empty$ \\
        singleton table $(k, v)$ & $k \mapsto \lbag v \rbag$ \\
        union of tables & $\finitemap{\keyset}{(\uplus)}\ \cdot\ merge$ \\
        projection $\projsymb{f}$ & $\finitemap{\keyset}{(\finitebag{f})}$ \\
        selection $\selectsymb{p}$ & $\finitemap{\keyset}{(filter\ p)}$ \\
        aggregation in monoid $\monoid{M}$ & $\finitemap{\keyset}{(reduce\
        \monoid{M})}$\\
            natural join & \finitemap{\keyset}{(\times)}\ $\cdot\ merge$ \\
    \end{tabular}
    \caption{Relational algebra operators implemented for bags}
    \label{tab:BagRelAlgOps}
\end{table}
