\section{Functions to benchmark}\label{sec:benchmark:functions}
In this section we present an overview of the three functions we will
be comparing in this benchmark. The paper \relalg{} presents a solution for
equijoin (and arbitrary selections followed by joins) query optimisation using
structures in Haskell, the algorithms compared will comprise of two alternative
``dummy'' solutions which may be seen as a simple implementation without the
findings of the paper, and a solution using an indexed structure as described
in the paper.

\paragraph{Product equijoin} The first function to be presented is
$productEquijoin$. This function
expresses the modular nature of the equijoin query (each part is expressed in
discrete steps) while using a non-indexed approach resulting in a global
Cartesian product.
The function has two discrete steps: first it must compute the product of the
two relations and then it must filter the result based on the equijoin criteria.
This is achieved by the following code.

\input{code/productEquijoin}

\noindent
The two discrete steps are clearly visible and separated by the composition of
functions. The first step is to compute the Cartesian product using the
$cp$ function. As a reminder the $cp$ function has
the following type signature.
\input{code/cpTypeSignature}

\noindent
The $cp$ function works as any Cartesian product would and simply takes a pair of bags and
returns a bag constituting the pairs of elements their elements.
The second step is to select the elements from the pair of
bags. The type signature of $select$ is as follows.
\input{code/selectTypeSignature}

\noindent
As you can see, $select$ is simply a
$filter$ on bags. The complexity of the function is housed in the
equality function. The equality function is modified to transform the bags using
the given functions before comparison. This is to generalise the idea of
comparing certain attributes; as mentioned in \fref{sec:relationalmodel} an
equijoin is a join based on equality of certain attributes and therefore just as
a record would need a transformation to compare attribute, we arbitrarily
transform the bag.

\paragraph{Comprehension equijoin} The second of the `dummy' solutions is the
$comprehensionEquijoin$.
This solution also does not make use of the indexed structures implemented and
presented. It is
clear that the function just names a list comprehension. We expect this
comprehension to take an element from both bags and, if the predicate required
by the equijoin is satisfied, form a pair. The following code demonstrates
this.

\input{code/comprehensionEquijoin.tex}

\noindent
As is to be expected, the previous code takes the form of a monad comprehension.
This is possible using the GHC extension \verb|MonadComprehensions| and allowed by
the fact that $Bag$ is a monad. We see many similar
features to the $productEquijoin$ function. A major difference is that there is
no explicit product operations but the pair is created explicitly when matching
results are found (however, all possible pairs are still considered). The equality, though
not explicitly named as a helper function, also includes the transformation described above.

\paragraph{Indexed equijoin} The final function is the only of the three to use
indexing. The indexing tested in this benchmark is basic however and further
optimisations are not considered. To this extent this function can also be seen as
modular; where a query engine or
custom written query could easily tune and combine steps, this function follows an
order putting together whole steps in a deterministic and segregated way.
This algorithm is by far
the most complex and is a version of what is proposed in the paper \relalg{}. In
simple terms, we index the bags and merge the result. After modification it to form
another valid indexed
structure we reduce this higher level construct back to a bag. The code for
these operations is as follows.

\input{code/indexedEquijoin}

\noindent
In order to understand the code it may be useful to explain some naming
conventions. Preceding $i$s are used to denote indexed or indexing structures,
for instance $if1$ is an `indexing function'. Furthermore, the letter $t$
refers to tables. As with previous functions it operates on two bags and
so the indices in variables refer to which bag the function is referring to.

As discussed above, the first step is to index the bags. This is achieved using
an inline form of the $indexBy$ function whose type
signature can be seen below. The $indexBy$ function
takes a table and a function that can transform the table to a key and indexes
the table by the key. The result of indexing a bag is simply a structure that
associates to each key a bag of the values that share that key. Now, with a pair of indexed tables
$merge$, whose type signature can be seen below,
combines the two indexed tables to one such that for each key a pair of bags is
associated to it. In order to complete reach a usable structure we must perform
a local Cartesian product (hence the $fmap cp$) on
each pair of bags to finally reach a single bag per key. Now we have a simple
indexed table again we can reduce it back to a bag (of pairs) using the reduce function
(whose type signature is below). More details on the structures used here can be
found in \fref{chap:database}.

\input{code/indexedEquijoinTypeSignatures}
