\chapter*{Acknowledgements}
\begin{comment}
It is usual to thank those individuals who have provided particularly useful assistance, technical or otherwise, during your project. Your supervisor will obviously be pleased to be acknowledged as they will have invested quite a lot of time into overseeing your progress and supporting you.
\end{comment}
\begin{center}
There are many people to acknowledge for their support during this project.

\noindent
\paragraph{} My supervisor, Dr.~Nicolas Wu, who not only inspired me to take the
amazing new passion of category theory but kept his supervision grounded in
reality and helped me make decisions that best impacted my project and the
context around it. Also my second marker, Dr~Steffen van Bakel who provided some
valuable advice during our meeting.

\noindent
\paragraph{} Bina Vishnuram, my study mentor, who has sat with me for countless
hours while I brainstormed, designed and wrote my project. She not only endured
my incoherent explanations of topics as I digested them but changed my
mind on what it means to succeed. It is no
overstatement that without her support this project and much of my future work
would not take the form it has and will.

\noindent
\paragraph{} My parents who have supported me, especially throughout the period
of this project. I appreciate all the extra work they took on for me to help me
concentrate on this project in periods of difficulty over the last year.

\noindent
\paragraph{} Mark Walker, Ibraheem Wazir and Elysia Williams who have provided
continuous support over the last year and helped me concentrate on this project;
both the project and the last year would not have been the same without them.
Also Sophie, Michael and any other friends who have supported me directly or
indirectly with their kind words and patience as my time was consumed by this
endeavour.
\end{center}
