\chapter*{Ethical Issues}
\begin{comment}
You are required to include a short discussion of ethical, legal, societal and professional issues that are relevant to your project (usually 1 to 2 pages long). This should fit in the most appropriate section of your project report, often the Background or Conclusions section.
\end{comment}
The project is an ethically neutral project. The domain of databases is
pervasive in the modern world and so the project is not likely to have an
especially biased impact on any particular group of people. Although language of
choice, Haskell, is occasionally used in safety critical environments, code
developed for the project is not likely to be used in such situations.

A large portion of the project revolves around generating data, however. This
could potentially pose a legal and privacy risk to individuals. Fortunately, the
final design is based purely on synthetic non-qualitative data and so does not
impact anybody. However, previous iterations did have a qualitative component to
it and there are definitely traces of such iterations in the codebase. When
collecting data to inform the synthesis, however, I used exclusively
information provided by the American government in public domain. Furthermore,
as it is distributed by the American government I am able to access the
information due to the Freedom of Information Act 2000. Usually more care would
have to be taken as the design and collation of data into a database is
protected by some forms of copyright.
